---
title: "Favorite foods"
author: "Olivia Zhang"
date: "`r Sys.Date()`"
output: 
  html_document: 
    toc: yes
    toc_float: yes
---

```{r load-packages}
library(tidyverse)
library(readxl)
library(stats)
```

# Part 1: Excel to CSV Workflow


## Step 1: Read the Data

* Read in the Excel file called `favourite-food.xlsx` from the `data-raw/` folder.

```{r read-in-data}
fav_food <- read_excel("data-raw/favourite-food.xlsx")
fav_food
```

## Step 2: Clean the Data

* Clean up the missing data (`NA`s) and make sure you're happy with variable types. Modify the `read_excel` function to take care of these issues.

```{r clean-na}
fav_food <- read_excel("data-raw/favourite-food.xlsx", 
                       na = c("N/A", "99999"))
fav_food <- fav_food %>%
  mutate(
    AGE = case_when(
      AGE == "five" ~ "5", 
      TRUE           ~ AGE
    ), 
    AGE = as.numeric(AGE)
  )

fav_food
```

* Convert the SES (socioeconomic status) to a factor variables with levels in the 
following order: `Low`, `Middle`, `High`.

```{r factor-ses}
fav_food %>%
  mutate(SES = fct_relevel(SES, "Low", "Middle", "High"))
```

## Step 3: Save to CSV

* Write out the resulting data frame to `favourite-food.csv` in the `data/` folder.

```{r write-favourite-food}
write_csv(fav_food, file = "data-raw/favourite-food.csv")

fs::dir_ls("data-raw")
```

## Step 5: Verify CSV Data

* Finally, read `favourite-food.csv` back in from the `data/` folder and 
observe the variable types. Are they as you left them?

```{r check-favourite-food}
favourite_food <- read.csv("data-raw/favourite-food.csv")
favourite_food
class(favourite_food$SES)
```

# Part 2: Excel to RDS Workflow

## Step 1: Read the Data
* Similar to Part 1, read  the Excel file called `favourite-food.xlsx` from the `data-raw/` folder, and handle missing data and variable types.

```{r read-in-data2}
fav_food <- read_excel("data-raw/favourite-food.xlsx", 
                       na = c("N/A", "99999"))
fav_food <- fav_food %>%
  mutate(
    AGE = case_when(
      AGE == "five" ~ "5", 
      TRUE           ~ AGE
    ), 
    AGE = as.numeric(AGE)
  )

fav_food
```

## Step 2: Adjust Variable Types

* Convert SES (socioeconomic status) to a factor variables with levels in the 
following order: `Low`, `Middle`, `High`.

```{r factor2}
fav_food %>%
  mutate(SES = fct_relevel(SES, "Low", "Middle", "High"))
```

## Step 3: Save to RDS
* Write out the resulting data frame to `favourite-food.rds` in the `data/` folder.

```{r write-rds}
write_rds(fav_food, file = "data-raw/favourite-food.rds")

fs::dir_ls("data-raw")
```

* Read `favourite-food.rds` back in from the `data/` folder and observe the 
variable types. Are they as you left them?

```{r label-read-rds}
favourite_food <- read_rds("data-raw/favourite-food.rds")
favourite_food
class(favourite_food$SES)
```

## References {#references}
1. Assignment Adapted from [Mine Cetinkaya-Rundel's Data Science in a Box](https://github.com/rstudio-education/datascience-box)